Dish: Distribution-oblivious Shell Scripting

### 2019-11-23 pldi/draft.tex

@@

Nodes $f$ of the graph represent functions from one (possibly empty)
input stream to an output stream $f : D^* \rightarrow D^*$. $D$
represents the basic data type of a line of characters. This
representation captures the majority of shell commands. The nodes are
monotonic, meaning that they cannot retract their output. More
formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there exists
$y'$ such that $f(x.x') = y.y'$, where $.$ is standard concatenation
and will be omitted when obvious. An example of a node is the
following command:

@@

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. This is the basic optimization
that is used by popular distributed processing systems such as
MapReduce\kk{cite} and Apache Spark\kk{cite}.Unfortunately, the
dataflow graph model of most of these systems is not expressive enough
to represent data parallelism, so the distributed implementations that
they perform do not preserve the semantics of their dataflow model.

@@

>>>>

\newcommand{\kstar}{^{\textstyle *}}
\newcommand{\eps}{\varepsilon}

@@

For a set $D$, we write $D\kstar$ to denote the set of all finite
words over $D$. For words $x, y \in D\kstar$, we write $x \cdot y$
or $xy$ to denote their concatenation. We write $\eps$ for the empty
word. We say that $x$ is a \emph{prefix} of $y$, and we write $x \leq y$, if there is a word $z$ such that $y = xz$. It is easy to see that $\leq$ is a partial order, and it is often referred to as the \emph{prefix order}.

A node $f$ of the graph represents a function from one (possibly empty)
input stream to an output stream $f : D\kstar \to D\kstar$, where $D$
represents the basic data type of a line of characters. This
representation captures the majority of shell commands. We require that
the function $f$ is monotone w.r.t.\ the prefix order. This captures
the idea that a node cannot retract output after it has been emitted.
%More formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there exists $y'$ such that $f(x.x') = y.y'$, where $.$ is standard
%concatenation and will be omitted when obvious.
An example of a node is the following command:

@@

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. This is the basic optimization
that is used by popular distributed processing systems such as
MapReduce \cite{DG2004MR} and Apache Spark \cite{spark:12}. Unfortunately, the
dataflow graph model of most of these systems is not expressive enough
to represent data parallelism \km{not sure this is true}, so the distributed implementations that
they perform do not preserve the semantics of their dataflow model.

@@

\km{Some very early references on the dataflow model of computation (maybe there are relevant?): \cite{KM1966, D1974Dataflow, K1974KPN, KMacQ1977}}

\cite{HSSGG2014, SHGW2015} discuss optimizing transformations, but their correctness is established using informal arguments based on operational intution.

The paper \cite{MSAIT2019} proposes a denotational semantic framework for stream processing where the data streams are seen as partial orders, and establishes the soundness of some common parallelizing transformations on dataflow graphs.
